Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}

### 2021-01-25 main.tex
Minor fixes

%\begin{example}
  Figure~\ref{fig:example-decompose} shows an example of a decomposed derivation in $\FOKS$ of the
  formula $(\exists x. \cneg{p}(x)) \vlor (\forall y.(p(y) \vlan
  p(f(y))))$.
%\end{example}

\begin{figure}
  \begin{equation*}
    \scalebox{.9}{$
    \vlderivation{
      \vlin{\acD}{}{(\exists x. \cneg{p}x) \vlor (\forall y.(py \vlan pfy))}{
        \vlin{\mexD}{}{(\exists x. (\cneg{p}x \vlor \cneg{p}x) \vlor (\forall
          y.(py \vlan pfy))}{
          \vlin{\fequ}{}{((\exists x. \cneg{p}x) \vlor (\exists x.\cneg{p}x))
            \vlor (\forall y.(py \vlan pfy)}{
            \vlin{\exists}{}{\forall y.(((\exists x. \cneg{p}x) \vlor (\exists
              x.\cneg{p}x)) \vlor (py \vlan pfy))}{
              \vlin{\exists}{}{\forall y.((\cneg{p}y \vlor (\exists x.\cneg{p}x))
                \vlor (py \vlan pfy))}{
                \vlin{\fequ}{}{\forall y.((\cneg{p}y \vlor \cneg{p}fy) \vlor
                  (py \vlan pfy))}{
              \vlin{\sw}{}{\forall y.(\cneg{p}y \vlor ((py \vlan pfy)
                \vlor \cneg{p}fy))}{
                \vlin{\fequ}{}{\forall y.(\cneg{p}y \vlor (py \vlan (pfy
                  \vlor \cneg{p}fy)))}{
                  \vlin{\aiD}{}{\forall y.((\cneg{p}y \vlor py) \vlan (pfy
                    \vlor \cneg{p}fy))}{
                    \vlin{\aiD}{}{\forall y.((\cneg{p}y \vlor py) \vlan
                      \ttt)}{
                      \vlin{\ttt}{}{\forall y.(\ttt \vlan \ttt)}{
                        \vlin{\forall}{}{\forall y.\ttt}{
                          \vlhy{\ttt}}}}}}}}}}}}}}
    $}
  \end{equation*}    
    \caption{Example derivation in decomposed form of Theorem~\ref{thm:decomposition}}
    \label{fig:example-decompose}
  \end{figure}
%
There is a weaker version of Theorem~\ref{thm:decomposition} that will
also be useful:
%
\begin{thm}\label{thm:decompositionA}
  For every derivation $\upsmash{\vlder{\FOKS}{\Deri}{A}{\ttt}}$ there is a formula~$A_1$ not containing any occurence of $\fff$, such that there is a derivation:
  \begin{equation*}
    %\scalebox{.9}{$
    \vlderivation{
      \vlde{\set{\wrD,\cD,\fequ}}{}{A}{
        \vlde{\FOMLS}{}{A_1}{
          \vlhy{\ttt}}}}
   % $}
  \end{equation*}
\end{thm}

@@

\begin{thm}\label{thm:CP-DI}
  Let  $\phi\colon\gC\to\gA$ be a combinatorial proof and let $A$ be a formula with $\gA=\graphof A$. Then there is a derivation
  \begin{equation}
    \label{eq:decom}
    \vlderivation{
      \vlde{\set{\wrD,\wfaD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri_2}{A}{
        \vlde{\FOMLS}{\Deri_1}{A'}{
          \vlhy{\ttt}}}}
  \end{equation}
  for some $A'\fequ C\rsubstof\phi$ where $C$ is a formula with $\graphof
  C=\gC$ and $\rsubstof\phi$ is the variable renaming substitution
  induced by~$\phi$.  Conversely,
  whenever we have a derivation as in~\eqref{eq:decom} above, such that $\fff$ does not occur in~$A_1$, then
  there is a combinatorial proof $\phi\colon\gC\to\graphof A$ such
  that $\gC=\graphof{\rectif{A'}}$.
\end{thm}

@@

\begin{proof}
  We only show the cases for $\wfaD$ and $\cfaD$:
  $$
  \todo{}
  \vlderivation{
    \vlin{\cD}{}{\forall x.A}{
      \vlin{\fequ}{(x \notin fv(A))}{\forall x.(A \vlor A)}{
        \vlin{\wrD}{}{A \vlor (\forall x.A)}{
          \vlhy{A}}}}}
  $$.
  $$
  \vlderivation{
    \vlin{\cD}{}{\forall x.A}{
      \vlin{\equiv}{(x \notin fv(\forall x.A))}{(\forall x.A) \vlor (\forall x.A)}{
        \vlin{\wrD}{}{\forall x.((\forall x.A) \vlor A)}{
          \vlhy{\forall x.\forall x.A}}}}}
  $$
  The other cases have been shown before (e.g.~\cite{brunnler:phd}).
%%   We have the following derivations:
%%   $$
%%   \vlderivation{
%%     \vlin{\cD}{}{\forall x.A}{
%%       \vlin{\equiv}{(x \notin fv(\forall x.A))}{(\forall x.A) \vlor (\forall x.A)}{
%%         \vlin{\wrD}{}{\forall x.((\forall x.A) \vlor A)}{
%%           \vlhy{\forall x.\forall x.A}}}}}
%%   $$.
%%   $$
%%   \vlderivation{
%%     \vlin{\cD}{}{(A \vlor B) \vlan (C \vlor D)}{
%%       \vlin{\fequ}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((A \vlor B) \vlan (C \vlor D))}{
%%       \vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((B \vlor A) \vlan (D \vlor C))}{
%% 	\vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((B \vlor A) \vlan D))}{
%% 	  \vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor (B \vlan D))}{
%% 	    \vlin{\wrD}{}{((A \vlor B) \vlan C) \vlor (B \vlan D)}{
%% 	      \vlhy{(A \vlan C) \vlor (B \vlan D)}}}}}}}}
%%   $$.
%%   $$
%%   \vlderivation{
%%     \vlin{\cD}{}{\exists x.(A \vlor B)}{
%%       \vlin{\fequ}{}{\exists x.(A \vlor B)) \vlor (\exists x.(A \vlor B))}{
%%       \vlin{\wrD}{}{(\exists x.(A \vlor B)) \vlor (\exists x.(B \vlor A))}{
%% 	\vlin{\wrD}{}{(\exists x.(A \vlor B)) \vlor (\exists x.B)}{
%% 	  \vlhy{(\exists x.A) \vlor (\exists x.B)}}}}}}
%%   $$.
%%   $$
%%   \vlderivation{
%%     \vlin{\cD}{}{\forall x.(A \vlor B)}{
%%       \vlin{\fequ}{}{(\forall x.(A \vlor B)) \vlor (\forall x.(A \vlor B))}{
%%       \vlin{\wrD}{}{(\forall x.(A \vlor B)) \vlor (\forall x.(B \vlor A))}{
%% 	\vlin{\wrD}{}{(\forall x.(A \vlor B)) \vlor (\forall x.B)}{
%% 	  \vlhy{(\forall x.A) \vlor (\forall x.B)}}}}}}
%%   $$.
%%   \juihsuan{done. If needed, we can introduce the notion of open deduction to
%% reduce the size of derivations...}\lutz{I was thinking about that, but (i) it is probably not worth the effort, as we wont have many derivation, and (ii) it is hard to define rectified derivations this way.}
\end{proof}

@@

\begin{proof}[Proof of Theorem~\ref{thm:MLS1->fonet}]
  From $\Deri$ we construct a derivation $\frameof\Deri$ of $\frameof
  A$ in the propositional fragment of $\FOMLS$, such that
  \hbox{$\fographof{\frameof\Deri}=\frameof{\fographof\Deri}$}.  The
  rules $\aiD,\tttD,\mix$ and $\sw$ are translated trivially, and for
  $\fequ$, it suffices to observe that the frame construction is
  invariant under $\fequ$. Finally, for the rules $\forall$ and
  $\exists$, proceed as follows. Every instance of $\forall$ is
  replaced by the derivation on the right below:\footnote{For better
  readability we omit superfluous parentheses, knowing that we always
  have $\fequ$ incorporating associativity and commutativity of
  $\vlan$ and $\vlor$.}
  \begin{equation*}\small
    \vlinf{\forall}{}{\Scons{\forall y_j.\ttt}}{\Scons{\ttt}}
    %\qquad
    \;\leadsto
    %\qquad
    \vlderivation{
      \vlde{\set{\sw,\fequ}}{\DDeri_2}{\Scons{q_{h_1,j}\vlor\cdots\vlor
          q_{h_j,j}\vlor(\dual q_{h_1,j}\vlan\cdots\vlan\dual
          q_{h_j,j}\vlan\ttt)}}{
        \vlde{\set{\aiD,\tttD}}{\DDeri_1}{
          \Scons{(q_{h_1,j}\vlor \dual q_{h_1,j})\vlan\cdots\vlan(q_{h_j,j}\vlor \dual q_{h_j,j})\vlan\ttt}}{
          \vlhy{\ttt}}}}
  \end{equation*}
  where $h_1,\ldots,h_j$ range over the indices of the existential
  binders dependend on that $y_j$. It is easy to see how $\DDeri_1$ is
  constructed, and for $\DDeri_2$ see,
  e.g. \cite{ESSLI19-notes,gug:str:01,brunnler:tiu:01}\lutz{check if
    it is really there. otherwise \cite{dissvonlutz}} Then, every
  occurrence of $\forall y_j.F$ is replaced by
  $q_{h_1,j}\vlor\cdots\vlor q_{h_j,j}\vlor (\dual
  q_{h_1,j}\vlan\cdots\vlan\dual q_{h_j,j}\vlan F)$ in the derivation
  below that $\forall$-instance. Now, observe that all instances of
  the $\exists$-rule introducing $x_i$ dependend on $y_j$ must occur
  below in the derivation (otherwise $\Deri$ would not be recified).
  Now consider such an instance $\vlinf{\exists}{}{\Scons{\exists
      x_i.B}}{\Scons{B\ssubst {x_i} t}}$.  Its context $\Sconhole$
  must contain all the $\forall y_j$ the $\exists x_i$ depends on,
  such that $B$ is in their scope. Following the translation of the
  $\forall$ rules above, we can therefore translate the $\exists$-rule
  instance by the following derivation
  \begin{equation*}
    \vlder{\set{\sw,\fequ}}{\DDeri_3}{ S_0\cons{S_1\cons{\cdots
          S_{k_i-1}\cons{S_{k_i}\cons{\dual q_{i,k_1}\vlan\dual
              q_{i,k_2}\vlan\cdots q_{i,k_i}\vlan B'}}\cdots}}}{
      S_0\cons{\dual q_{i,k_1}\vlan S_1\cons{\dual
          q_{i,k_2}\vlan\cdots S_{k_i-1}\cons{\dual q_{i,k_i}\vlan
            S_{k_i}\cons{B'}}\cdots}}}
  \end{equation*}
  where $k_1,\ldots,k_i$ are the indices of of the universal binders
  on which that $x_i$ depends, and $B'$ is $B$ in which all predicates
  are replaced by nullary one (step~3 in the frame construction). The
  derivation $\DDeri_3$ can be constructed in the same way as
  $\DDeri_2$ above.

@@

\begin{lemma}\label{lem:cw->rectif}
  Let $\downsmash{\vlderivation{
    \vlde{\set{\wrD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri}{B}{
      \vlhy{A}}}}$ be a derivation. Then there is a recified derivation
  $\vlderivation{
    \vlde{\set{\wrD,\rectif\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif\Deri}{\rectif
      B}{ \vlhy{\rectif A}}}$, such that the induced maps
  $\mapof\Deri\colon\graphof{A}\to\graphof{B}$ and
  $\mapof{\rectif\Deri}\colon\graphof{\rectif A}\to\graphof{\rectif
    B}$ are equal up to a variable renaming of the vertex labels.
\end{lemma}

@@

  For showing that it is also a skew fibration, we construct for
  $\Deri$ its propositional encoding $\PE\Deri$ by
  translating every line into its propositional encoding.
  \juihsuan{maybe mention that an instance of one of the other rules can be translated
into an instance of the same rule. It's trivial but may be worth mentioning.}\lutz{done below}
  The instances
  of the rules $\rectif\mfaD$ and $\rectif\mexD$ are replaced in two
  steps by:
  \begin{equation*}
    \vlderivation{
      \vlin{\rectif\acD}{}{\Scons{x\vlor(\PE{(Ax)}\vlor\PE{(Bx)})}}{
        \vlin{\fequ}{}{\Scons{(y\vlor z)\vlor(\PE{(Ay)}\vlor\PE{(Bz)})}}{
          \vlhy{\Scons{(y\vlor\PE{(Ay)})\vlor (z\vlor\PE{(Bz)})}}}}}
  \end{equation*}
  and
  \begin{equation*}
    \vlderivation{
      \vlin{\rectif\acD}{}{\Scons{x\vlan(\PE{(Ax)}\vlor\PE{(Bx)})}}{
        \vlin{\me}{}{\Scons{(y\vlor z)\vlan(\PE{(Ay)}\vlor\PE{(Bz)})}}{
          \vlhy{\Scons{(y\vlan\PE{(Ay)})\vlor (z\vlan\PE{(Bz)})}}}}}
  \end{equation*}
  respectively, where $\rectif\acD$ is a $\acD$ that renames the
  variables---the propositional variable, as well as the first-order
  variable of the same name---as everything is recified, there is no
  ambiguity here. Any instance of a rule $\wD$, $\acD$, $\me$, or
  $\fequ$ is translated to an instance of the same rule, and
  $\rectif\cfaD$ is translated to $\rectif\acD$.

@@

  Conversely, assume we have a decomposed derivation
  \begin{equation}
    \label{eq:decomx}
    \vlderivation{
      \vlde{\set{\wrD,\acD,\me,\mfaD,\mexD,\fequ}}{\Deri_2}{A}{
        \vlde{\FOMLS}{\Deri_1}{A'}{
          \vlhy{\ttt}}}}
  \end{equation}
  Then we can transform $\Deri_1$ into a recified form
  $\rectif\Deri_1$, proving $\rectif A'$. By
  Theorem~\ref{thm:MLS1->fonet}, the linked fograph
  $\fographof{\rectif\Deri_1}=\tuple{\fographof{\rectif
      A'},\linkingof{\rectif\Deri_1}}$ is a fonet.  Then, by
  Lemma~\ref{lem:cw->rectif}, there is a rectified derivation
  $\vlderivation{
    \vlde{\set{\wrD,\rectif\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif{\Deri_2}}{\rectif
      A}{ \vlhy{\rectif{A'}}}}$ whose induced map
  $\mapof{\rectif{\Deri_2}}\colon\graphof{\rectif{A'}}\to\graphof{\rectif
    A}$ is the same a the induced map
  $\mapof{\Deri_2}\colon\graphof{A'}\to\graphof{A}$ of $\Deri_2$. By
  Lemma~\ref{lem:cw->skew}, this map is a skew bifibration. Hence, we
  have a combinatorial proof $\phi\colon\gC\to\graphof{A}$ with
  $\gC=\graphof{\rectif{A'}}$.
  \lutz{shit, something's wrong...}
  %% Let $\downsmash{\vlderivation{
  %%   \vlde{\set{\wrD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri}{B}{
  %%     \vlhy{A}}}}$ be a derivation. Then there is a recified derivation
  %% $\vlderivation{
  %%   \vlde{\set{\wrD,\rectif\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif\Deri}{\rectif
  %%     B}{ \vlhy{\rectif A}}}$, such that the induced maps
  %% $\mapof\Deri\colon\graphof{A}\to\graphof{B}$ and
  %% $\mapof{\rectif\Deri}\colon\graphof{\rectif A}\to\graphof{\rectif
  %%   B}$ are equal up to a variable renaming of the vertex labels.
  %% for some $A'\fequ C\sigma$ where $C$ is a formula with $\graphof
  %% C=\gC$ and $\sigma$ is a variable renaming substitution.  Conversely,
  %% whenever we have a derivation as in~\eqref{eq:decom} above, then
  %% there is a combinatorial proof $\phi\colon\gC\to\graphof A$ such
  %% that $\gC=\graphof{\rectif{A'}}$.
\end{proof}

>>>>

%\begin{example}
  Figure~\ref{fig:example-decompose} shows an example of a decomposed derivation in $\FOKS$ of the
  formula $(\exists x. \cneg{p}x) \vlor (\forall y.(py \vlan pfy))$.
%\end{example}

\begin{figure}
  \begin{equation*}
    \scalebox{.9}{$
    \vlderivation{
      \vlin{\acD}{}{(\exists x. \cneg{p}x) \vlor (\forall y.(py \vlan pfy))}{
        \vlin{\mexD}{}{(\exists x. (\cneg{p}x \vlor \cneg{p}x) \vlor (\forall
          y.(py \vlan pfy))}{
          \vlin{\fequ}{}{((\exists x. \cneg{p}x) \vlor (\exists x.\cneg{p}x))
            \vlor (\forall y.(py \vlan pfy)}{
            \vlin{\exists}{}{\forall y.(((\exists x. \cneg{p}x) \vlor (\exists
              x.\cneg{p}x)) \vlor (py \vlan pfy))}{
              \vlin{\exists}{}{\forall y.((\cneg{p}y \vlor (\exists x.\cneg{p}x))
                \vlor (py \vlan pfy))}{
                \vlin{\fequ}{}{\forall y.((\cneg{p}y \vlor \cneg{p}fy) \vlor
                  (py \vlan pfy))}{
              \vlin{\sw}{}{\forall y.(\cneg{p}y \vlor ((py \vlan pfy)
                \vlor \cneg{p}fy))}{
                \vlin{\fequ}{}{\forall y.(\cneg{p}y \vlor (py \vlan (pfy
                  \vlor \cneg{p}fy)))}{
                  \vlin{\aiD}{}{\forall y.((\cneg{p}y \vlor py) \vlan (pfy
                    \vlor \cneg{p}fy))}{
                    \vlin{\aiD}{}{\forall y.((\cneg{p}y \vlor py) \vlan
                      \ttt)}{
                      \vlin{\ttt}{}{\forall y.(\ttt \vlan \ttt)}{
                        \vlin{\forall}{}{\forall y.\ttt}{
                          \vlhy{\ttt}}}}}}}}}}}}}}
    $}
  \end{equation*}    
    \caption{Example derivation in decomposed form of Theorem~\ref{thm:decomposition}}
    \label{fig:example-decompose}
  \end{figure}
%
There is a weaker version of Theorem~\ref{thm:decomposition} that will
also be useful:
%
\begin{thm}\label{thm:decompositionA}
  For every derivation $\upsmash{\vlder{\FOKS}{\Deri}{A}{\ttt}}$ there is a formula~$A_1$ not containing any occurrence of $\fff$, such that there is a derivation:
  \begin{equation*}
    %\scalebox{.9}{$
    \vlderivation{
      \vlde{\set{\wrD,\cD,\fequ}}{}{A}{
        \vlde{\FOMLS}{}{A_1}{
          \vlhy{\ttt}}}}
   % $}
  \end{equation*}
\end{thm}

@@

\begin{thm}\label{thm:CP-DI}
  Let  $\phi\colon\gC\to\gA$ be a combinatorial proof and let $A$ be a formula with $\gA=\graphof A$. Then there is a derivation
  \begin{equation}
    \label{eq:decom}
    \vlderivation{
      \vlde{\set{\wrD,\wfaD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri_2}{A}{
        \vlde{\FOMLS}{\Deri_1}{A'}{
          \vlhy{\ttt}}}}
  \end{equation}
  for some $A'\fequ C\rsubstof\phi$ where $C$ is a formula with $\graphof
  C=\gC$ and $\rsubstof\phi$ is the variable renaming substitution
  induced by~$\phi$.  Conversely,
  whenever we have a derivation as in~\eqref{eq:decom} above, such that $\fff$ does not occur in~$A'$, then
  there is a combinatorial proof $\phi\colon\gC\to\graphof A$ such
  that $\gC=\graphof{\rectif{A'}}$.
\end{thm}

@@

\begin{proof}
  We only show the cases for $\wfaD$ and $\cfaD$:
  $$
  \vlderivation{
    \vlin{\cD}{}{\forall x.A}{
      \vlin{\fequ}{(x \notin fv(A))}{\forall x.(A \vlor A)}{
        \vlin{\wrD}{}{A \vlor (\forall x.A)}{
          \vlhy{A}}}}}
  $$.
  $$
  \vlderivation{
    \vlin{\cD}{}{\forall x.A}{
      \vlin{\equiv}{(x \notin fv(\forall x.A))}{(\forall x.A) \vlor (\forall x.A)}{
        \vlin{\wrD}{}{\forall x.((\forall x.A) \vlor A)}{
          \vlhy{\forall x.\forall x.A}}}}}
  $$
  The other cases have been shown before (e.g.~\cite{brunnler:phd}).
%%   We have the following derivations:
%%   $$
%%   \vlderivation{
%%     \vlin{\cD}{}{\forall x.A}{
%%       \vlin{\equiv}{(x \notin fv(\forall x.A))}{(\forall x.A) \vlor (\forall x.A)}{
%%         \vlin{\wrD}{}{\forall x.((\forall x.A) \vlor A)}{
%%           \vlhy{\forall x.\forall x.A}}}}}
%%   $$.
%%   $$
%%   \vlderivation{
%%     \vlin{\cD}{}{(A \vlor B) \vlan (C \vlor D)}{
%%       \vlin{\fequ}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((A \vlor B) \vlan (C \vlor D))}{
%%       \vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((B \vlor A) \vlan (D \vlor C))}{
%% 	\vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((B \vlor A) \vlan D))}{
%% 	  \vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor (B \vlan D))}{
%% 	    \vlin{\wrD}{}{((A \vlor B) \vlan C) \vlor (B \vlan D)}{
%% 	      \vlhy{(A \vlan C) \vlor (B \vlan D)}}}}}}}}
%%   $$.
%%   $$
%%   \vlderivation{
%%     \vlin{\cD}{}{\exists x.(A \vlor B)}{
%%       \vlin{\fequ}{}{\exists x.(A \vlor B)) \vlor (\exists x.(A \vlor B))}{
%%       \vlin{\wrD}{}{(\exists x.(A \vlor B)) \vlor (\exists x.(B \vlor A))}{
%% 	\vlin{\wrD}{}{(\exists x.(A \vlor B)) \vlor (\exists x.B)}{
%% 	  \vlhy{(\exists x.A) \vlor (\exists x.B)}}}}}}
%%   $$.
%%   $$
%%   \vlderivation{
%%     \vlin{\cD}{}{\forall x.(A \vlor B)}{
%%       \vlin{\fequ}{}{(\forall x.(A \vlor B)) \vlor (\forall x.(A \vlor B))}{
%%       \vlin{\wrD}{}{(\forall x.(A \vlor B)) \vlor (\forall x.(B \vlor A))}{
%% 	\vlin{\wrD}{}{(\forall x.(A \vlor B)) \vlor (\forall x.B)}{
%% 	  \vlhy{(\forall x.A) \vlor (\forall x.B)}}}}}}
%%   $$.
%%   \juihsuan{done. If needed, we can introduce the notion of open deduction to
%% reduce the size of derivations...}\lutz{I was thinking about that, but (i) it is probably not worth the effort, as we wont have many derivation, and (ii) it is hard to define rectified derivations this way.}
\end{proof}

@@

\begin{proof}[Proof of Theorem~\ref{thm:MLS1->fonet}]
  From $\Deri$ we construct a derivation $\frameof\Deri$ of $\frameof
  A$ in the propositional fragment of $\FOMLS$, such that
  \hbox{$\fographof{\frameof\Deri}=\frameof{\fographof\Deri}$}.  The
  rules $\aiD,\tttD,\mix$ and $\sw$ are translated trivially, and for
  $\fequ$, it suffices to observe that the frame construction is
  invariant under $\fequ$. Finally, for the rules $\forall$ and
  $\exists$, proceed as follows. Every instance of $\forall$ is
  replaced by the derivation on the right below:\footnote{For better
  readability we omit superfluous parentheses, knowing that we always
  have $\fequ$ incorporating associativity and commutativity of
  $\vlan$ and $\vlor$.}
  \begin{equation*}\small
    \vlinf{\forall}{}{\Scons{\forall y_j.\ttt}}{\Scons{\ttt}}
    %\qquad
    \;\leadsto
    %\qquad
    \vlderivation{
      \vlde{\set{\sw,\fequ}}{\DDeri_2}{\Scons{q_{h_1,j}\vlor\cdots\vlor
          q_{h_j,j}\vlor(\dual q_{h_1,j}\vlan\cdots\vlan\dual
          q_{h_j,j}\vlan\ttt)}}{
        \vlde{\set{\aiD,\tttD}}{\DDeri_1}{
          \Scons{(q_{h_1,j}\vlor \dual q_{h_1,j})\vlan\cdots\vlan(q_{h_j,j}\vlor \dual q_{h_j,j})\vlan\ttt}}{
          \vlhy{\ttt}}}}
  \end{equation*}
  where $h_1,\ldots,h_j$ range over the indices of the existential
  binders dependend on that $y_j$. It is easy to see how $\DDeri_1$ is
  constructed, and for $\DDeri_2$ see,
  e.g. \cite{ESSLI19-notes,gug:str:01,brunnler:tiu:01}\lutz{check if
    it is really there. otherwise \cite{dissvonlutz}} Then, every
  occurrence of $\forall y_j.F$ is replaced by
  $q_{h_1,j}\vlor\cdots\vlor q_{h_j,j}\vlor (\dual
  q_{h_1,j}\vlan\cdots\vlan\dual q_{h_j,j}\vlan F)$ in the derivation
  below that $\forall$-instance. Now, observe that all instances of
  the $\exists$-rule introducing $x_i$ dependend on $y_j$ must occur
  below in the derivation (otherwise $\Deri$ would not be rectified).
  Now consider such an instance $\vlinf{\exists}{}{\Scons{\exists
      x_i.B}}{\Scons{B\ssubst {x_i} t}}$.  Its context $\Sconhole$
  must contain all the $\forall y_j$ the $\exists x_i$ depends on,
  such that $B$ is in their scope. Following the translation of the
  $\forall$ rules above, we can therefore translate the $\exists$-rule
  instance by the following derivation
  \begin{equation*}
    \vlder{\set{\sw,\fequ}}{\DDeri_3}{ S_0\cons{S_1\cons{\cdots
          S_{k_i-1}\cons{S_{k_i}\cons{\dual q_{i,k_1}\vlan\dual
              q_{i,k_2}\vlan\cdots q_{i,k_i}\vlan B'}}\cdots}}}{
      S_0\cons{\dual q_{i,k_1}\vlan S_1\cons{\dual
          q_{i,k_2}\vlan\cdots S_{k_i-1}\cons{\dual q_{i,k_i}\vlan
            S_{k_i}\cons{B'}}\cdots}}}
  \end{equation*}
  where $k_1,\ldots,k_i$ are the indices of of the universal binders
  on which that $x_i$ depends, and $B'$ is $B$ in which all predicates
  are replaced by nullary one (step~3 in the frame construction). The
  derivation $\DDeri_3$ can be constructed in the same way as
  $\DDeri_2$ above.

@@

\begin{lemma}\label{lem:cw->rectif}
  Let $\downsmash{\vlderivation{
    \vlde{\set{\wrD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri}{B}{
      \vlhy{A}}}}$ be a derivation. Then there is a rectified derivation
  $\vlderivation{
    \vlde{\set{\wrD,\rectif\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif\Deri}{\rectif
      B}{ \vlhy{\rectif A}}}$, such that the induced maps
  $\mapof\Deri\colon\graphof{A}\to\graphof{B}$ and
  $\mapof{\rectif\Deri}\colon\graphof{\rectif A}\to\graphof{\rectif
    B}$ are equal up to a variable renaming of the vertex labels.
\end{lemma}

@@

  For showing that it is also a skew fibration, we construct for
  $\Deri$ its propositional encoding $\PE\Deri$ by
  translating every line into its propositional encoding.
  \juihsuan{maybe mention that an instance of one of the other rules can be translated
into an instance of the same rule. It's trivial but may be worth mentioning.}\lutz{done below}
  The instances
  of the rules $\rectif\mfaD$ and $\rectif\mexD$ are replaced in two
  steps by:
  \begin{equation*}
    \vlderivation{
      \vlin{\rectif\acD}{}{\Scons{x\vlor(\PE{(Ax)}\vlor\PE{(Bx)})}}{
        \vlin{\fequ}{}{\Scons{(y\vlor z)\vlor(\PE{(Ay)}\vlor\PE{(Bz)})}}{
          \vlhy{\Scons{(y\vlor\PE{(Ay)})\vlor (z\vlor\PE{(Bz)})}}}}}
  \end{equation*}
  and
  \begin{equation*}
    \vlderivation{
      \vlin{\rectif\acD}{}{\Scons{x\vlan(\PE{(Ax)}\vlor\PE{(Bx)})}}{
        \vlin{\me}{}{\Scons{(y\vlor z)\vlan(\PE{(Ay)}\vlor\PE{(Bz)})}}{
          \vlhy{\Scons{(y\vlan\PE{(Ay)})\vlor (z\vlan\PE{(Bz)})}}}}}
  \end{equation*}
  respectively, where $\rectif\acD$ is a $\acD$ that renames the
  variables---the propositional variable, as well as the first-order
  variable of the same name---as everything is rectified, there is no
  ambiguity here. Any instance of a rule $\wD$, $\acD$, $\me$, or
  $\fequ$ is translated to an instance of the same rule, and
  $\rectif\cfaD$ is translated to $\rectif\acD$.

@@

  Conversely, assume we have a decomposed derivation
  \begin{equation}
    \label{eq:decomx}
    \vlderivation{
      \vlde{\set{\wrD,\acD,\me,\mfaD,\mexD,\fequ}}{\Deri_2}{A}{
        \vlde{\FOMLS}{\Deri_1}{A'}{
          \vlhy{\ttt}}}}
  \end{equation}
  Then we can transform $\Deri_1$ into a rectified form
  $\rectif\Deri_1$, proving $\rectif A'$. By
  Theorem~\ref{thm:MLS1->fonet}, the linked fograph
  $\fographof{\rectif\Deri_1}=\tuple{\fographof{\rectif
      A'},\linkingof{\rectif\Deri_1}}$ is a fonet.  Then, by
  Lemma~\ref{lem:cw->rectif}, there is a rectified derivation
  $\vlderivation{
    \vlde{\set{\wrD,\rectif\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif{\Deri_2}}{\rectif
      A}{ \vlhy{\rectif{A'}}}}$ whose induced map
  $\mapof{\rectif{\Deri_2}}\colon\graphof{\rectif{A'}}\to\graphof{\rectif
    A}$ is the same a the induced map
  $\mapof{\Deri_2}\colon\graphof{A'}\to\graphof{A}$ of $\Deri_2$. By
  Lemma~\ref{lem:cw->skew}, this map is a skew bifibration. Hence, we
  have a combinatorial proof $\phi\colon\gC\to\graphof{A}$ with
  $\gC=\graphof{\rectif{A'}}$.
  \lutz{shit, something's wrong...}
  %% Let $\downsmash{\vlderivation{
  %%   \vlde{\set{\wrD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri}{B}{
  %%     \vlhy{A}}}}$ be a derivation. Then there is a recified derivation
  %% $\vlderivation{
  %%   \vlde{\set{\wrD,\rectif\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif\Deri}{\rectif
  %%     B}{ \vlhy{\rectif A}}}$, such that the induced maps
  %% $\mapof\Deri\colon\graphof{A}\to\graphof{B}$ and
  %% $\mapof{\rectif\Deri}\colon\graphof{\rectif A}\to\graphof{\rectif
  %%   B}$ are equal up to a variable renaming of the vertex labels.
  %% for some $A'\fequ C\sigma$ where $C$ is a formula with $\graphof
  %% C=\gC$ and $\sigma$ is a variable renaming substitution.  Conversely,
  %% whenever we have a derivation as in~\eqref{eq:decom} above, then
  %% there is a combinatorial proof $\phi\colon\gC\to\graphof A$ such
  %% that $\gC=\graphof{\rectif{A'}}$.
\end{proof}
